Search, Compress, Compile: Library Learning in Neurally-Guided Bayesian Program Learning

### 2018-05-18 nips_2018.tex

We trained our system on a corpus of 109 automatically generated text editing tasks, with 4 input/output examples each.
After three iterations, it assembles a DSL containing a dozen new functions (center of Fig.~\ref{initialExampleDSL}) that let it solve 
all of the training tasks.
But, how well does the  learned DSL generalized to real text-editing scenarios?
We tested, but did not train, on the 108 text editing problems from the SyGuS~\cite{alur2016sygus} program synthesis competition. Before any learning,
\system solves 3.7\% of the problems with an average search time of 235 seconds.
After learning,
it solves 74.1\%, and does so much faster,
solving them in an average of 29 seconds.
As of the 2017 SyGuS competition,
the best-performing algorithm solves 79.6\% of the problems.
But, SyGuS comes with a
different hand-engineered DSL \emph{for each text editing problem}.\footnote{SyGuS text editing problems also prespecify the set of allowed string constants \emph{for each task}. Our system does not use this assistance.}
Here  we learned a single DSL
that applied generically to
all of the tasks,
and perform comparably to the best
prior work.

@@

We compare with ablations of our model on held out tasks.
The purpose of this ablation study is 
both to examine the role of each component of \systemEnding,
as well as to compare with
prior approaches in the literature:
a head-to-head
comparison of program synthesizers is complicated by the fact that
each system, including ours, makes idiosyncratic 
assumptions about the space of programs and the statement of tasks.
\begin{wraptable}{r}{0.55\textwidth}
\vspace{-5pt}
%% This radically shrinks the table
\tabcolsep=2.5pt
\renewcommand{\arraystretch}{0.5}
\begin{tabular}{lcccccc}
  \toprule& Ours& 
    No NN & SE&NPS & PCFG & Enum \\
  \midrule
  \multicolumn{7}{c}{\emph{List Processing}}\\
  \midrule
  \% solved&\textbf{79\%}\footnote{A last-minute bug caused this  to run without ``Helmholtz machine'' style training ($\mathcal{L}_{\text{RM}} = \mathcal{L}_{\text{SS}}$). We are rerunning this experiment and expect this number to be higher.} &
  76\% &71\%&35\%&62\%&37\%\\
  Solve time&  4.1s&5.8s&10.6s&34.7s&43.4s&20.2s\\
  \midrule
  \multicolumn{7}{c}{\emph{Text Editing}}\\
  \midrule
  \% solved&\textbf{74\%} &43\% &30\%&33\%&0\%&4\%\\
  Solve time& 29s&65s&38s&80s&--&235s\\
  \midrule
  \multicolumn{7}{c}{\emph{Symbolic Regression}}\\
  \midrule
  \% solved&   \textbf{84\% }&75\%&62\%&38\%&38\%&37\% \\
  Solve time&  24s& 40s  &28s&31s&55s&29s\\
  \bottomrule
  \end{tabular}
\caption{\% solved before timeout. Solve time: averaged over solved
  tasks.  %SE: NPS: trained like RobustFill/DeepCoder.  PCFG: model w/o
  %structure learning.  Enum: model w/o any learning.  %% MDL:
  %% $-\expect\left[\probability[x|p] \right]$. For domains other than
  %% symbolic regression MDL is 0 nats.
  %% For symbolic regression MDL is a
  %% proxy for \# of continuous parameters.
}\vspace{-0.5cm}\label{baselineComparisons} \end{wraptable}
Nevertheless, much prior work can be modeled within our setup. 
We compare with the following ablations (Table~\ref{baselineComparisons}):
\\\noindent \textbf{No NN:} lesions the recognition model.
\\\noindent \textbf{NPS}, which does not learn the DSL,
instead learning the recognition model
from samples drawn from the fixed DSL.
We call this NPS (Neural Program Synthesis)
because this is closest to how
RobustFill~\cite{devlin2017robustfill} and DeepCoder~\cite{balog2016deepcoder} are trained.
\\\noindent \textbf{SE}, which lesions the recognition model and restricts the DSL  learning algorithm to
only add \textbf{s}ub\textbf{e}xpressions of programs in the frontiers to the DSL. This is how prior approaches have learned libraries of functions~\cite{Dechter:2013:BLV:2540128.2540316,DBLP:conf/icml/LiangJK10,DBLP:conf/ecai/LinDETM14}.
\\\noindent \textbf{PCFG}, which lesions the recognition model and does not learn the DSL,
but instead learns the parameters of the DSL ($\theta$) -- learning the parameters of a PCFG while not learning any of the structure.
\\\noindent \textbf{Enum}, which enumerates a frontier without any learning -- equivalently, our first search step.

%For each domain,
%% We are interested both in how many tasks the
%% agent can solve and how quickly it can find those solutions.
%% Tbl.~\ref{baselineComparisons}
%% compares our model against these alternatives.
%% We consistently
%% improve on the baselines,
%% and find that lesioning the recognition model
%% and lesioning it also slows down the convergence of the algorithm,
%% taking more iterations to reach a given number of tasks solved (Fig.~\ref{learningCurves}).
%% This supports a view of the recognition model as a way of amortizing the cost of search.

>>>>

We trained our system on a corpus of 109 automatically generated text editing tasks, with 4 input/output examples each.
After three iterations, it assembles a DSL containing a dozen new functions (center of Fig.~\ref{initialExampleDSL}) that let it solve 
all of the training tasks.
But, how well does the  learned DSL generalized to real text-editing scenarios?
We tested, but did not train, on the 108 text editing problems from the SyGuS~\cite{alur2016sygus} program synthesis competition. Before any learning,
\system solves 3.7\% of the problems with an average search time of 235 seconds.
After learning,
it solves 74.1\%, and does so much faster,
solving them in an average of 29 seconds.
As of the 2017 SyGuS competition,
the best-performing algorithm solves 79.6\% of the problems.
But, SyGuS comes with a
different hand-engineered DSL \emph{for each text editing problem}.\footnote{SyGuS text editing problems also prespecify the set of allowed string constants for each task. For these experiments, our system did not use this assistance.}
Here  we learned a single DSL
that applied generically to
all of the tasks,
and perform comparably to the best
prior work.

@@

We compare with ablations of our model on held out tasks.
The purpose of this ablation study is 
both to examine the role of each component of \systemEnding,
as well as to compare with
prior approaches in the literature:
a head-to-head
comparison of program synthesizers is complicated by the fact that
each system, including ours, makes idiosyncratic 
assumptions about the space of programs and the statement of tasks.
\begin{wraptable}{r}{0.55\textwidth}
\vspace{-5pt}
%% This radically shrinks the table
\tabcolsep=2.5pt
\renewcommand{\arraystretch}{0.5}
\begin{tabular}{lcccccc}
  \toprule& Ours& 
    No NN & SE&NPS & PCFG & Enum \\
  \midrule
  \multicolumn{7}{c}{\emph{List Processing}}\\
  \midrule
  \% solved&\textbf{79\%}\footnotemark & 
  76\% &71\%&35\%&62\%&37\%\\
  Solve time&  4.1s&5.8s&10.6s&34.7s&43.4s&20.2s\\
  \midrule
  \multicolumn{7}{c}{\emph{Text Editing}}\\
  \midrule
  \% solved&\textbf{74\%} &43\% &30\%&33\%&0\%&4\%\\
  Solve time& 29s&65s&38s&80s&--&235s\\
  \midrule
  \multicolumn{7}{c}{\emph{Symbolic Regression}}\\
  \midrule
  \% solved&   \textbf{84\% }&75\%&62\%&38\%&38\%&37\% \\
  Solve time&  24s& 40s  &28s&31s&55s&29s\\
  \bottomrule
  \end{tabular}
\caption{\% solved before timeout. Solve time: averaged over solved
  tasks.  %SE: NPS: trained like RobustFill/DeepCoder.  PCFG: model w/o
  %structure learning.  Enum: model w/o any learning.  %% MDL:
  %% $-\expect\left[\probability[x|p] \right]$. For domains other than
  %% symbolic regression MDL is 0 nats.
  %% For symbolic regression MDL is a
  %% proxy for \# of continuous parameters.
}\vspace{-0.5cm}\label{baselineComparisons} \end{wraptable}
Nevertheless, much prior work can be modeled within our setup. 
We compare with the following ablations (Table~\ref{baselineComparisons}):
\\\noindent \textbf{No NN:} lesions the recognition model.
\\\noindent \textbf{NPS}, which does not learn the DSL,
instead learning the recognition model
from samples drawn from the fixed DSL.
We call this NPS (Neural Program Synthesis)
because this is closest to how
RobustFill~\cite{devlin2017robustfill} and DeepCoder~\cite{balog2016deepcoder} are trained.
\\\noindent \textbf{SE}, which lesions the recognition model and restricts the DSL  learning algorithm to
only add \textbf{S}ub\textbf{E}xpressions of programs in the frontiers to the DSL. This is how most prior approaches have learned libraries of functions~\cite{Dechter:2013:BLV:2540128.2540316,DBLP:conf/icml/LiangJK10,DBLP:conf/ecai/LinDETM14}.
\\\noindent \textbf{PCFG}, which lesions the recognition model and does not learn the DSL,
but instead learns the parameters of the DSL ($\theta$) -- learning the parameters of a PCFG while not learning any of the structure.
\\\noindent \textbf{Enum}, which enumerates a frontier without any learning -- equivalently, our first search step.
\footnotetext[2]{Due to a last-minute bug we ran the list processing experiments without training the recognition model on ``Helmholtz machine'' style data. We are rerunning this experiment and anticipate improved quantitative results for list processing -- notice the recognition model helps much more for text editing and symbolic regression.} &
%For each domain,
%% We are interested both in how many tasks the
%% agent can solve and how quickly it can find those solutions.
%% Tbl.~\ref{baselineComparisons}
%% compares our model against these alternatives.
%% We consistently
%% improve on the baselines,
%% and find that lesioning the recognition model
%% and lesioning it also slows down the convergence of the algorithm,
%% taking more iterations to reach a given number of tasks solved (Fig.~\ref{learningCurves}).
%% This supports a view of the recognition model as a way of amortizing the cost of search.
